---
title: "Sales Mehod Review"
author: "Sarim Rizvi"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
theme_set(theme_minimal())
```

## Project Overview

This project analyzes the efficiency of three sales strategies — **Email**, **Call**, and **Email + Call** — using real-world product sales data from a new office stationery line. The goal is to recommend the most scalable, cost-effective method based on revenue and effort..

## Data Summary

```{r Load Data}
sales_clean <- read_csv("data/product_sales_clean.csv")
glimpse(sales_clean)
```
- Raw data had information about 150000 customers.

- Removed 1074 rows due to missing revenue, and 2 rows due to invalid customer tenure.

- Total rows after cleaning: 13924

## Exploratory Data Analysis
### Sales Method Distribution
```{r Sales Method Plot}
sales_clean %>%
  count(sales_method) %>%
  ggplot(aes(x = sales_method, y = n, fill = sales_method)) +
  geom_col() +
  labs(title = "Customers by Sales Method", x = "Method", y = "Count") +
  theme(legend.position = "none")
```

## Revenue Spread over Sales Method
```{r Revenue 1}
ggplot(sales_clean, aes(x = sales_method, y = revenue, fill = sales_method)) +
  geom_boxplot() +
  labs(title = "Revenue per Customer by Method", x = "Method", y = "Revenue ($)") +
  theme(legend.position = "none")
```


## Revenue Over Time across Sales Method
```{r Revenue 2}
sales_clean %>%
  group_by(week, sales_method) %>%
  summarise(avg_revenue = mean(revenue), .groups = "drop") %>%
  ggplot(aes(x = week, y = avg_revenue, color = sales_method)) +
  geom_line(size = 1.2) +
  labs(title = "Average Revenue Over Time by Method", y = "Avg Revenue ($)", color = 'Sales Method')
```

## Metric: Revenue per Minute
```{r Metrics}
sales_metrics <- sales_clean %>%
  mutate(
    estimated_minutes = case_when(
      sales_method == "Email" ~ 1,
      sales_method == "Call" ~ 30,
      sales_method == "Email + Call" ~ 11
    ),
    revenue_per_unit = revenue / nb_sold,
    revenue_per_minute = revenue / estimated_minutes
  )

metric_summary <- sales_metrics %>%
  group_by(sales_method) %>%
  summarise(
    avg_revenue = mean(revenue),
    avg_units = mean(nb_sold),
    avg_rev_per_unit = mean(revenue_per_unit),
    avg_rev_per_min = mean(revenue_per_minute),
    avg_effort = mean(estimated_minutes),
    customers = n(),
    .groups = "drop"
  )

print(metric_summary)
```

## Key Insights
- **Email** yields high revenue per minute (~$97), best for scale.

- **Email + Call** brings the highest revenue per customer (~$184), ideal for high-value leads.

- **Call-only** performs poorly in both revenue and efficiency.

## Recommendation
- Use **Email** as **default method** for all campaigns.

- Use **Email + Call** for **top-tier accounts** based on past spending.

- Avoid Call-only unless targeting niche, high-stakes clients.

- Monitor **revenue per minute** weekly by method.


